Covariate analysis of ROSMAP RNASeq

### 2016-01-27 code/Rmd/MSSM_RNASeq.Rmd
Minor modifications

# Find significant covariates
adjustCovars = designMatVarsToCovars(getCovariatesMap(DM),preAdjustedSigCovars$significantCovars)
```
Significant covariates to adjust at FDR 0.1 are `r adjustCovars`
```{r preAdjustedSigCovars.NULL, fig.width=15, fig.height=8}
preAdjustedSigCovars[["PC_res"]][[2]]$plotData
```
### Normalisation (custom/manual design)
Re-perform normalisation with custom design matrix (manually with "TISSUE.Dx","SEX","RACE","PMI" as fixed effects and Donor as random effects)
```{r reperform.normalisation, cache=TRUE}
# Post adjusted design matrix
DM1 = getDesignMatrix(COVARIATES[,c("TISSUE.Dx","SEX","RACE","PMI"),drop=F],Intercept = F)
DM1$design = DM1$design[,linColumnFinder(DM1$design)$indepCols]

@@

```{r residualSigCovars.manual, fig.width=15, fig.height=8}
residualSigCovars[["PC_res"]][[2]]$plotData
```
### Residual calculation
Calculate weighted residuals and add back "TISSUE.Dx" to the residuals
```{r varsToAddBack}
# Add variable of interest back to the residuals
varsToAddIn = c("TISSUE.DxBM_10.AD", "TISSUE.DxBM_10.Control", "TISSUE.DxBM_22.AD",
                "TISSUE.DxBM_22.Control", "TISSUE.DxBM_36.AD","TISSUE.DxBM_36.Control")
RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION + 
  VOOM.ADJUSTED.FIT$coefficients[,varsToAddIn] %*% t(DM1$design[,varsToAddIn])
```

@@

# Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"),
                       filters = "ensembl_gene_id", values = rownames(RESIDUAL.GENE_EXPRESSION),
                       mart = mart)
RESIDUAL.GENE_EXPRESSION <- left_join(rownameToFirstColumn(RESIDUAL.GENE_EXPRESSION, "ensembl_gene_id"),
                                      Ensemble2HGNC) %>%
  dplyr::select(ensembl_gene_id, hgnc_symbol, one_of(rownames(COVARIATES)))
```
```{r summary, results='asis', echo=FALSE}
writeLines('Number of samples in each category are')
tmp = as.data.frame(summary(COVARIATES$TISSUE.Dx))
colnames(tmp) = 'No. of samples'
kable(tmp)
```
### Differential expression analysis
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diffExp}
# Fit contrast
contrast = makeContrasts(contrasts=c("TISSUE.DxBM_10.Control-TISSUE.DxBM_10.AD",
                                     "TISSUE.DxBM_22.Control-TISSUE.DxBM_22.AD",
                                     "TISSUE.DxBM_36.Control-TISSUE.DxBM_36.AD"), 
                         levels = colnames(VOOM.ADJUSTED.FIT$coefficients))
FIT = contrasts.fit(VOOM.ADJUSTED.FIT, contrasts=contrast)
FIT = eBayes(FIT)

>>>>

# Find significant covariates
adjustCovars = designMatVarsToCovars(getCovariatesMap(DM),preAdjustedSigCovars$significantCovars)
```
Significant covariates to adjust at FDR 0.1 are `r adjustCovars`
```{r preAdjustedSigCovars.NULL, fig.width=20, fig.height=12}
preAdjustedSigCovars[["PC_res"]][[2]]$plotData
```
### Normalisation (custom/manual design)
Re-perform normalisation with custom design matrix (manually with "TISSUE.Dx","SEX","RACE","PMI" as fixed effects and Donor as random effects)
```{r reperform.normalisation, cache=TRUE}
# Post adjusted design matrix
DM1 = getDesignMatrix(COVARIATES[,c("TISSUE.Dx","SEX","RACE","PMI"),drop=F],Intercept = F)
DM1$design = DM1$design[,linColumnFinder(DM1$design)$indepCols]

@@

```{r residualSigCovars.manual, fig.width=20, fig.height=12}
residualSigCovars[["PC_res"]][[2]]$plotData
```
### Residual calculation
Calculate weighted residuals and add back "TISSUE.Dx" to the residuals
```{r varsToAddBack}
# Add variable of interest back to the residuals
varsToAddIn = c("TISSUE.DxBM_10.AD", "TISSUE.DxBM_10.Control", "TISSUE.DxBM_22.AD",
                "TISSUE.DxBM_22.Control", "TISSUE.DxBM_36.AD","TISSUE.DxBM_36.Control")
RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION + 
  VOOM.ADJUSTED.FIT$coefficients[,varsToAddIn] %*% t(DM1$design[,varsToAddIn])
```

@@

# Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"),
                       filters = "ensembl_gene_id", values = rownames(RESIDUAL.GENE_EXPRESSION),
                       mart = mart)
RESIDUAL.GENE_EXPRESSION <- left_join(rownameToFirstColumn(RESIDUAL.GENE_EXPRESSION, "ensembl_gene_id"),
                                      Ensemble2HGNC) %>%
  dplyr::select(ensembl_gene_id, hgnc_symbol, one_of(rownames(COVARIATES)))
```
```{r summary, results='asis', echo=FALSE}
writeLines('Number of samples in each category are')
tmp = as.data.frame(summary(COVARIATES$TISSUE.Dx))
colnames(tmp) = 'No. of samples'
kable(tmp)
```
### Differential expression analysis
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diffExp, fig.height=10, fig.width=18}
# Fit contrast
contrast = makeContrasts(contrasts=c("TISSUE.DxBM_10.Control-TISSUE.DxBM_10.AD",
                                     "TISSUE.DxBM_22.Control-TISSUE.DxBM_22.AD",
                                     "TISSUE.DxBM_36.Control-TISSUE.DxBM_36.AD"), 
                         levels = colnames(VOOM.ADJUSTED.FIT$coefficients))
FIT = contrasts.fit(VOOM.ADJUSTED.FIT, contrasts=contrast)
FIT = eBayes(FIT)

### 2016-01-27 code/Rmd/ROSMAP_RNASeq.Rmd
Minor modifications

# Store covariates
write.table(RESIDUAL.GENE_EXPRESSION, file = 'expression.tsv', sep = '\t', row.names=F, quote=F)
EXP_OBJ = File('expression.tsv', name = 'FPKM Normalised Adjusted Residual Expression (cogdx added)', parentId = CODE$properties$id)
EXP_OBJ = synStore(EXP_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                   executed = thisFile, activityDescription = activityDescription)
```
|  *Results*            |  *SynapseID*                    |
|  -------              |   ---------                     |
|  Covariates           |  `r COV_OBJ$properties$id`      |
|  Residual Expression  |  `r EXP_OBJ$properties$id`      |

>>>>

# Store covariates
write.table(RESIDUAL.GENE_EXPRESSION, file = 'expression.tsv', sep = '\t', row.names=F, quote=F)
EXP_OBJ = File('expression.tsv', name = 'FPKM Normalised Adjusted Residual Expression (cogdx added)', parentId = CODE$properties$id)
EXP_OBJ = synStore(EXP_OBJ, used = ALL_USED_IDs, activityName = activityName, 
                   executed = thisFile, activityDescription = activityDescription)

```
|  *Results*            |  *SynapseID*                    |
|  -------              |   ---------                     |
|  Covariates           |  `r COV_OBJ$properties$id`      |
|  Residual Expression  |  `r EXP_OBJ$properties$id`      |
